%!TEX root = PhD_Thesis.tex
\chapter{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%======================================================================
\section{Motivation}
%======================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Needles are slim metal tubes with sharpened tips that allow access to internal anatomy with minimal damage to the skin and other intervening tissues. Needles are extremely ubiquitous tools in modern medicine. They are applied in a range of form factors by clinicians in many specialties to inject drugs, drain fluids, remove tissue samples for testing, or deposit implants. Interventional radiology is a medical specialty that uses instruments inserted through needles under medical image guidance to treat disease. Such percutaneous interventions are increasingly favored as alternatives to surgery, as they result in less trauma to the patient, and thus reduce the risk of complications and recovery time. 

Liver tumors, percutaneous ablation.

Robotic needle steering

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%======================================================================
\section{Contributions}
%======================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We briefly summarize the major contributions of this dissertation as follows:
\begin{itemize}
\item We did some stuff.
\item We did some more stuff.
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%======================================================================
\section{Prior Work}
%======================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Techniques for Needle Steering}


\subsection{Automatic Segmentation of Needles from Ultrasound Data}
\cite{Adebar2013}
\cite{Adebar2015}

\subsection{Control Approaches for Robotic Needle Steering}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%======================================================================
\section{Dissertation Overview}
%======================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This thesis consists of six chapters. In chapter 1, which is this introduction, we presented the motivation for our research of using tactile feedback for sensory substitution and augmentation of force feedback. We also presented prior work on the various types of tactile feedback devices and prior work on providing interaction force information in robot-assisted minimally invasive surgery.

Chapter 2 describes doppler imaging...

Chapter 3 describes needles...

Chapter 4 describes teleoperation...

Chapter 5 summarizes the results of this research, reviews the contribution made in this dissertation, and provides suggestions for future work.